Natural transformations

### 2017-07-21 papers/base-3/paper-3.tex
merged with Maxi

\newcounter{counter}[section]
\newtheorem{defn}[counter]{Definition}
\newtheorem{exmp}[counter]{Example}

@@

>>>>

\newcounter{counter}[section]
\newtheorem{defn}[counter]{Definition}
\newtheorem{exmp}[counter]{Example}
\newtheorem{lemma}[counter]{Lemma}

@@

\section{Equivalence of Categories}
\begin{exmp}
  \[
    \begin{tikzcd}
      \catA: & & \catB:\\
      A \arrow[loop left] & & B \arrow[bend left]{d} \arrow[loop left] \\
                          & & B' \arrow[bend left]{u} \arrow[loop left]
    \end{tikzcd}
  \]
\end{exmp}
%todo explain example

\begin{defn}
  \label{def:eqv}
  Tow categories $\catA$ and $\catB$ are \emph{equivalent} if:
  \begin{itemize}
    \item There are functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ and
    \item there are natural transformations $\eta \from 1_{\catA} \to G \circ F$ and $\varepsilon \from F \circ G \to 1_\catB$.
  \end{itemize}
  The functors $F$ and $G$ are called \emph{equivalences}.
  We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
\end{defn}

\begin{defn}
  \label{def:surj-on-obj}
  A functor $F \from \catA \to \catB$ is \emph{essentially surjective on objects}
  if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
\end{defn}

\begin{lemma}
  \label{lem:equiv}
  A functor $F \from \catA \to \catB$ is an equivalence if and only if $F$ is full, faithful and surjective on objects.
\end{lemma}

\begin{exmp}
  Let $K$ be a field.
  We define the category $\catA$ such that objects are of the form $K^n$ for $n \in \NN$ and morphisms are linear functions.
  Moreover we denote the category of finite dimensional vector spaces by $\FDVect_K$.
  Objects in this category are finite dimensional vector spaces and morphisms are linear functions.
  Let $F \from \catA \to \FDVect_K$ be the embedding functor from $\catA$ into $\FDVect_K$, i.e. $F$ maps objects and morphisms to themselves.
  Clearly $F$ is full and faithful.
  For surjectivity on objects we fix an arbitrary object $V \in \FDVect_K$.
  Let $n \in \NN$ be the dimension of $V$.
  From linear algebra we know that $V \iso K^n = F(K^n)$.
  Thus $F$ is essentially surjectivity on objects.
  By lemma~\ref{lem:equiv} we obtain that $\catA \eqv \FDVect_K$.
  %In order to show that $\catA \eqv \FDVect_K$ we have to define a functor $F \from \catA \to \FDVect_K$ which is full, faithful and surjective on objects.
\end{exmp}

\section{Horizontal Composition}

\[
\xymatrix@C+.5em{
\cat{A} \rtwocell<4>^F_G{\alpha}   &
\cat{A}' \rtwocell<4>^{F'}_{G'}{\alpha'}   &
\cat{A}''
}
\]
\[
\xymatrix@C+.5em{
  \cat{A} \rtwocell<4>^{F' \of F}_{G' \of G} &\cat{A}'',
}
\]

This horizontal composition is written $\alpha' * \alpha$.
The following diagram commutes because $\alpha'$ is a natural transformation.
\[
  \begin{tikzcd}
    F'(F(A)) \arrow{r}{F'(\alpha_A)} \arrow{d}{\alpha'_{F(A)}} \arrow{dr}{(\alpha' * \alpha)_A} & F'(G(A)) \arrow{d}{\alpha'_{G(A)}} \\
    G'(F(A)) \arrow{r}{G'(\alpha_A)}                                                            & G'(G(A))
  \end{tikzcd}
\]
So we can define $(\alpha' * \alpha)_A$ either as $G'(\alpha_A) \circ \alpha'_{F(A)}$ or as $\alpha'_{G(A)} \circ F'(\alpha_A)$.
Now we want to show that $\alpha' * \alpha$ is again a natural transformation.
Therefore let $f\from A \to A'$ be a morphism in $\catA$.
\[
  \begin{tikzcd}
    F'(F(A)) \arrow{r}{F'(F(f))} \arrow{d}{\alpha'_{F(A)}} & F'(F(A')) \arrow{d}{\alpha'_{F(A')}} \\
    G'(F(A)) \arrow{r}{G'(F(f))} \arrow{d}{G'(\alpha_A)}   & G'(F(A')) \arrow{d}{G'(\alpha_{A'}}  \\
    G'(G(A)) \arrow{r}{G'(G(f))}                           & G'(G(A'))
  \end{tikzcd}
\]

%todo example

We can also combine horizontal and vertical composition.
\[
\xymatrix{
\cat{A}
\ruppertwocell<8>^F{\alpha}
\ar[r]|G
\rlowertwocell<-8>_H{\beta} &
\cat{A}'
\ruppertwocell<8>^{F'}{\hspace{.2em}\alpha'}
\ar[r]|{G'}
\rlowertwocell<-8>_{H'}{\hspace{.2em}\beta'} &
\cat{A}''
}
\]

%todo interchange law
